Sperm should evolve to make female meiosis fair.

### 2014-12-22 writeup_and_notes/evolution/RESUBMISSIONmale_dependent_drive.tex
In blue, YB makes minor changes to GC improvements

It is therefore somewhat surprising that despite the intense evolutionary pressure on female meiosis to prevent meiotic drive, 
it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
That is, in many animal species, female meiosis is completed only
after fertilization \citep{Masui_book}, creating ample opportunity for interaction between the sperm and
	female meiotic machinery (note that the alternation of generations
	likely precludes this interaction in plants). 
\gc{The timing of enty of sperm into the egg relative to the
  progression of female meiosis is variable across animals, as is the
  stage at which female meiosis
  stalls without fertilization (we show the phylogenetic distribution
  in supplementary Figure XXX).}
Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females 
     could presumably rapidly spread through a population. 
At first sight, female meiosis appears primed for conflict caused by such selfish systems. 
Here we ask if sperm do indeed evolve to collaborate with female drivers to exploit this 
	apparent weakness in the defense against meiotic drive. \newline

@@

Inclusion of a heterozygous fitness cost (i.e. $w_{AB}=1-s_h$)
	further constrains the evolution of a self-promoting driver. 
In fact, with any heterozygous fitness cost, a rare self-promoting
	driver is always selected against. 
However, this case also displays bistability -- 
	when $s$ is sufficiently small (Equation \ref{sForHetFix}) this allele fixes deterministically if its 
	frequency exceeds some threshold  (Equation \ref{thresh2}, exact results in 
	Figure \ref{bistable_additive}).
This bistability prevents self-promoting drivers from invading 	
	reasonably sized populations, and assures that if they do invade, they will rapidly fix.
Our model therefore predicts that self-promoting drivers will not be
	observed as stable polymorphisms in natural populations. 
This lack of a balanced polymorphism %(Figure \ref{Invasion_space}) 
	precludes the evolution of an
	allele that suppresses this form of meiotic drive in females. 
\gc{Allowing for arbitrary levels of inbreeding (in the form of
  self-fertilization), as opposed to the panmictic model explored
  above -- more thoroughly aligns the interests of both parents and
  parental chromosomes, restricting further the possibility for
  invasion of both traditional female drivers and 'self-promoting'
  drivers (Figures SXX-SXY). Additionally, because inbreeding reduces
  the frequency of heterozygotes, the invasion and fixation criteria
  converge as both become more strict with selfing promoting rates.} \newline

>>>>

It is therefore somewhat surprising that despite the intense evolutionary pressure on female meiosis to prevent meiotic drive, 
it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
That is, in many animal species, female meiosis is completed only
after fertilization \citep{Masui_book}, creating ample opportunity for interaction between the sperm and
%	female meiotic machinery (note that the alternation of generations
%	likely precludes this interaction in plants). 
\yb{	female meiotic machinery (note that, across animals the variation in timing of sperm entry into the egg and the timing at which female meiosis stalls [supplementary Figure XXX] complicates this opportunity in some taxa, and that the alternation of generations likely precludes this interaction in plants).}
%\gc{The timing of entry of sperm into the egg relative to the
%  progression of female meiosis is variable across animals, as is the
%  stage at which female meiosis
%  stalls without fertilization (we show the phylogenetic distribution
%  in supplementary Figure XXX).}
Therefore, in many species a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in heterozygous females could presumably rapidly spread through a population. 
At first sight, female meiosis appears primed for conflict caused by such selfish systems. 
Here we ask if sperm do indeed evolve to collaborate with female drivers to exploit this 
	apparent weakness in the defense against meiotic drive. \newline

@@

Inclusion of a heterozygous fitness cost (i.e. $w_{AB}=1-s_h$)
	further constrains the evolution of a self-promoting driver. 
In fact, with any heterozygous fitness cost, a rare self-promoting
	driver is always selected against. 
However, this case also displays bistability -- 
	when $s$ is sufficiently small (Equation \ref{sForHetFix}) this allele fixes deterministically if its 
	frequency exceeds some threshold  (Equation \ref{thresh2}, exact results in 
	Figure \ref{bistable_additive}).
This bistability prevents self-promoting drivers from invading 	
	reasonably sized populations, and assures that if they do invade, they will rapidly fix.
Our model therefore predicts that self-promoting drivers will not be
	observed as stable polymorphisms in natural populations. 
This lack of a balanced polymorphism %(Figure \ref{Invasion_space}) 
	precludes the evolution of an
	allele that suppresses this form of meiotic drive in females. 
\gc{ \yb{Relaxing our assumptions of panmixia by} allowing for 
  arbitrary levels of inbreeding (in the form of self-fertilization),
  %, as opposed to the panmictic model exploredabove -- 
  more thoroughly aligns the interests of both parents and
  parental chromosomes, restricting further the possibility for
  invasion of both traditional female drivers and 'self-promoting'
  drivers (Figures SXX-SXY). Additionally, because inbreeding reduces
  the frequency of heterozygotes, the invasion and fixation criteria
  converge, \yb{as both become stricter with increased inbreeding rates}.} \newline
